Algebraic Groups\vspace{-1ex}

### 2020-01-27 Algebraic Geometry/Algebraic-Geometry.tex
Update

>>>>

\section{January 27th, 2020}

\subsection{Absolute properties of schemes}
\begin{defn}
	A schemes is:
	\begin{enumerate}
		\item \textbf{connected} (resp. \textbf{irreducible}) if the topological space $|X|$ is.
		\item \textbf{reduced} if for all $U\subseteq X$ $\O_X(U)$ is reduced.
		\item \textbf{integral} if for all $U\subseteq X$ $\O_X(U)$ is an integral domain (iff $X$ is reduced and irreducible).
		\item \textbf{quasicompact} if $|X|$ is quasicompact.
		\item \textbf{locally Noetherian} if there is an open cover $\Spec A_i=U_i\subseteq X$ with $A_i$ Noetherian.
		\item \textbf{Noetherian} if $X$ is locally Noetherian and quasicompact (iff there is a finite Noetherian open cover).
		\item \textbf{normal} if there is an open cover $\Spec A_i=U_i\subseteq X$ with $A_i$ normal (integrally closed domain).
	\end{enumerate}
\end{defn}

A question: if $\Spec A$ is locally Noetherian, is $A$ Noetherian? The answer is yes! We need a sublemma:
\begin{lem}
	Suppose that $\Spec B\subseteq \Spec A$ is open with $f\in A$ such that $D(f)\subseteq\Spec B$. Then $D(f)=D(g)$
	where $g=f|_{\Spec B}\in\O(\Spec B)=B.$
\end{lem}
Then we may replace the $\Spec A_i$ by $\Spec A_{f_i}$ such that $A_{f_i}$ is Noetherian. But then $A\to \prod_1^n A_{f_i}$ is a faithfully flat 
extension of a Noetherian ring, so it follows from the following:
\begin{lem}
	Let $A\to B$ be faithfully flat. Then $B$ being Noetherian implies $A$ is.
\end{lem}
You can see this by tensoring up a chain $I_\bullet$ in $A$ to a sequence $I_\bullet\otimes B\cong (I_iB)_{i\in\bbN}$. This is a sequence in $B$.

\subsection{Examples}
Consider $f\in k[x,y]$. THen $\Spec k[x,y]/(f)$ is integral.
Consider $n$ copies of $\A1$ glued along $\A1\setminus 0$. This is also integral.

Think about $\Spec k[x,y]/(xy)$. Draw a picture. Is it reduced/ireducible/connected?

### 2020-01-27 Algebraic Groups/Algebraic-Groups.tex
Update

>>>>

\section{January 27. 2020}
We'll pick back up with proving that a finitely generated Hopf algebra in characterisitc zero is reduced.

\subsection{The result}
\begin{thm}
	Let $A$ be a finitely generated Hopf algebra over $k$ with $\ch k=0$. THen $A$ is reduced.
\end{thm}
\begin{prf}
	Let $Y\in I$ be a nilpotent with $y^m=0$. We want to show that $y=0$. It sufficies to show this for $m=2$.
	We claim that $y\in \cap_{n\ge 1}I^n$. To see this, suppose that $y\in I^n\setminus I^{n+1}$.
	then we can write 
	\[y=\sum_{\sum m_i=n}a_{\underline m}x^{\underline m}+I^{n+1}\]
	by the previous lemma. Then consider 
	\[y^2=p(x_1,\dots,x_\Gamma)+I^{2n+1}\]
	where $p$ is some monomial of degree $2n$ with some nonzero coefficient. Hence $y^2$ is nonzero mod $I^{2n+1}$, a contradiction!
	Hence $y\in\cap I^n$.

	We actually need that $k=\bar k$. Let $g:A\to k$ be a point in $G_A$. Then define the translation map
	\[T_g:A\xrightarrow{\Delta}A\otimes A\xrightarrow{g\otimes \id}k\otimes A\simeq A\]
	which is a map of algebras and is, in fact, invertible.

	Define $\frakm_g=\ker g$ (think this is the maximal corresponding to the point $g$ in the spectrum). Then we claim that $T_g(I)=\frakm_g$.
	If $y\in I$ then $y\in \frakm_G$ (since it's nilpotent it lies in every maximal). Thus 
	by the previous claim $y\in\cap_{n\ge 1} \frakm_g^n$ and by the Krull intersection theorem, $y=0$ in $A_{\frakm_g}$ for all $\frakm_g$,
	whence $y=0.$
\end{prf}

\subsection{Properties of \texorpdfstring{$\Omega_A$}{Omega A}}
\begin{itemize}
	\item $\Omega_{A\times B}\cong\Omega_A\times\Omega_B$
	\item Let $S$ be a multiplicative system in $A$. Then  
	\[\Omega_{S^{-1}A}\cong \Omega_A\otimes_A S^{-1}A\]
	and in particular if $A$ is an integral domain, let $K=\operatorname{Frac}(A).$ Then 
	\[\Omega_{K/k}\cong \Omega_A\otimes_A K.\]
	\item Let $L/k$ be a finitely generated seperable field extension. That is we have 
	\[L\supset E=k(x_1,\dots,x_\Gamma)\supset k\]
	where $L/E$ is finite. Then $\dim_L\Omega_{L/k}=\operatorname{trdeg}_k L$ and $dx_1,\dots,dx_\Gamma$ forma a basis for $\Omega_{L/k}$.
	\item Let $G$ be an algebraic group scheme. Maybe assume $k=\bar k$. Assume that $k[G]$ is an integral domain. Let $K=\operatorname{Frac}(k[G])$. Then 
	\[\rank\Omega_{k[G]}=\operatorname{trdeg}_k K\]
\end{itemize}
The last item merits some discussion:
\begin{prf}
	By the third item above, we get 
	\[\Omega_{K/k}=\Omega_{k[G]}\otimes_{k[G]}K\]
	so 
	\[\operatorname{trdeg}_k K=\dim_K\Omega_{K/k}=\rank \Omega_{k[G]}\]
\end{prf}
\begin{defn}
	Let $G$ be an algebraic group scheme over $k$. Then $G$ is smooth if 
	\[\dim G=\Omega_{k[G]}\]
	where $\dim G=\dim_{\text{Krull}} k[G]$.
\end{defn}
\begin{thm}\label{thm:smoothgrps}
	Let $G$ be an algebraic group scheme over $k$. Then if
	\[\bar k[G]=k[G]\otimes_k \bar k\]
	is reduced, $G$ is smooth.
\end{thm}
\begin{cor}
	In characteristic zero, all algebraic group schemes are smooth.
\end{cor}
\begin{rmk}
	Smoothness is independent of extending scalars (as the rank and dimension in the definition are), but 
	you do need to extend scalars enough so that the nilpotents aren't ``hiding''. Thus it suffices to only extend to 
	the seperable closure of $k$ or any perfect field $L/k$.
\end{rmk}
The following discussion and proof can be found in section 14 (and partially 6.6) of Waterhouse.
\begin{prf}[of thm.~\ref{thm:smoothgrps}]
	First extend scalars up to $\bar k$. Second recall (from section 6.6) that for $A=k[G]$, the following are equivalent:
	\begin{itemize}
		\item $\pi_0(A)$ is trivial
		\item $\Spec A$ is connected
		\item $\Spec A$ is irreducible
		\item $A/\operatorname{Nil}(A)$ is an integral domain.
	\end{itemize}
	The second two are actually a statement about algebraic geometry (recall the geometry doesn't see the nilradical).

	Assume that $k[G]$ is reduced. THen we can assume that $G=G^0$ is connected since 
	\[k[G]\simeq k[G^0]\times\cdots\times k[G^0]\]
	where we have $|\pi_0(k[G])|$ copies in the product. Then we use the first property in the last section.

	For $G^0$ connected, the general properties in the last section imply that $k[G^0]$ is irreducible implies that $k[G^0]/\operatorname{Nil}(k[G^0])$ is 
	an integral gdomain. By assumption, the nilradical is trivial, so $k[G^0]$ itself is an integral domain.

	Let $K=\operatorname{Frac}(k[G^0])$. Then
	\[\dim G^0=\dim k[G^0]\stackrel{\text{AG}}{=}\operatorname{trdeg}_kK=\rank \Omega_{k[G^0]}\]
	where the $AG$ referenced above is essentially the Noether normalization lemma.
\end{prf}
